---
title: "inClassCoding"
author: "Antonio Escallón"
date: "10/6/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r histogram_1, echo=FALSE}
require(palmerpenguins)
hist(penguins$body_mass_g)
```

```{r multi_panel_1, echo= FALSe}
par(mfrow = c(2, 2))
hist(
penguins$body_mass_g,
main = "Histogram of Body Mass",
xlab = "Body Mass (g)")
hist(
penguins$flipper_length_mm,
main = "Histogram of Flipper Length",
xlab = "Flipper Length (mm)")
boxplot(
body_mass_g ~ species, data = penguins,
main = "Body Mass and Species",
ylab = "Body Mass (g)"
)
boxplot(
flipper_length_mm ~ species, data = penguins,
main = "Flipper Length and Species",
ylab = "Flipper Length (mm)"
)
```


```{r multi_panel_2, echo = FALSE, fig.width = 7, fig.height = 10}
par(mfrow = c(2, 2))
hist(
penguins$body_mass_g,
main = "Histogram of Body Mass",
xlab = "Body Mass (g)")
hist(
penguins$flipper_length_mm,
main = "Histogram of Flipper Length",
xlab = "Flipper Length (mm)")
boxplot(
body_mass_g ~ species, data = penguins,
main = "Body Mass and Species",
ylab = "Body Mass (g)"
)
boxplot(
flipper_length_mm ~ species, data = penguins,
main = "Flipper Length and Species",
ylab = "Flipper Length (mm)"
)
```


# A New Section With Tabs {.tabset}

## Tab 1

Here's some text

## Tab 2

Here's some graphics:

```{r tab_graph 1, echo = FALSE}
boxplot(
  flipper_length_mm ~ species, data = penguins,
  main = "Flipper Length and Species",
  ylab = "Flipper Length (mm)"
)

```

## Tab 3

Here's a numbered list:

1. Item 1
1. Item 2
1. Item 3
1. **Item 4** (in bold)

# New 1st Level header (not there are no tabs)


## Subsection 1

This section is not in a tab

## Subsection 2

Still no tabs!

# Tab Section 2 {.tabset .tabset-pills}

This seciton uses a 'pill style' of tabs!

## Tab 1

Here's some text

## Tab 2

Here's some graphics:

```{r tab_graph 1, echo = FALSE}
boxplot(
  flipper_length_mm ~ species, data = penguins,
  main = "Flipper Length and Species",
  ylab = "Flipper Length (mm)"
)

```

## Tab 3

Here's a numbered list:

1. Item 1
1. Item 2
1. Item 3
1. **Item 4** (in bold)

